### 2021-08-17 Bulkplasmon.code.tex
Adding a file per tikz figure and including them as archive in the main tex file to not to disturb the tex lecture.

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Shape: Rectangle [id:dp5857108563848268] 
\draw [draw opacity=0][fill={rgb, 255:red, 198; green, 244; blue, 240
  } ,fill opacity=0.48 ] (148,66) -- (390.17,66) -- (390.17,269) --
(148,269) -- cycle ;
%Shape: Circle [id:dp3780902414948827] 
\draw [dash pattern={on 0.84pt off 2.51pt}] (217.04,166.52)
.. controls (217.04,135.56) and (242.14,110.46) .. (273.1,110.46)
.. controls (304.07,110.46) and (329.17,135.56) .. (329.17,166.52)
.. controls (329.17,197.48) and (304.07,222.58) .. (273.1,222.58)
.. controls (242.14,222.58) and (217.04,197.48) .. (217.04,166.52) --
cycle ;
%Shape: Polygon Curved [id:ds008835726195197124] 
\draw [fill={rgb, 255:red, 248; green, 231; blue, 28 } ,fill opacity=1
] (277.08,151.5) .. controls (296.08,166.5) and (293.08,169.5)
.. (269.08,167.5) .. controls (248.08,193.5) and (254.08,143.5)
.. (277.08,151.5) -- cycle ;
%Straight Lines [id:da7163027701373094] 
\draw [color={rgb, 255:red, 0; green, 13; blue, 255 } ,draw opacity=1
] (250.37,181.61) -- (233.46,195.22) ; \draw [shift={(231.9,196.48)},
  rotate = 321.15999999999997] [color={rgb, 255:red, 0; green, 13;
    blue, 255 } ,draw opacity=1 ][line width=0.75] (6.56,-1.97)
.. controls (4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls
(1.99,0.18) and (4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da27099612271830587] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (256.82,188.49) -- (247.21,198.4) ; \draw [shift={(245.82,199.84)},
  rotate = 314.11] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da882164603438807] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(290.5,185.75) -- (306.35,200.58) ; \draw [shift={(307.81,201.95)},
  rotate = 223.09] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da884690742053359] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (296.41,178.39) -- (307.57,186.52) ; \draw [shift={(309.19,187.7)},
  rotate = 216.04] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da20074876184087243] 
\draw [color={rgb, 255:red, 41; green, 51; blue, 241 } ,draw opacity=1
] (270.3,191.49) -- (270.4,213.19) ; \draw [shift={(270.41,215.19)},
  rotate = 269.72] [color={rgb, 255:red, 41; green, 51; blue, 241 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da608342686522136] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (279.7,190.73) -- (281.46,204.42) ; \draw [shift={(281.72,206.41)},
  rotate = 262.67] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da088624174223913] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(247.34,162.12) -- (225.65,161.27) ; \draw [shift={(223.65,161.19)},
  rotate = 362.25] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da3058655897728151] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (247.69,171.55) -- (233.92,172.7) ; \draw [shift={(231.93,172.87)},
  rotate = 355.2] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da7992511633354091] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(296.59,151.57) -- (313.23,137.63) ; \draw [shift={(314.76,136.34)},
  rotate = 500.02] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da021188728886464392] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (290,144.82) -- (299.42,134.72) ; \draw [shift={(300.78,133.26)},
  rotate = 492.97] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da5888482352950647] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(254.98,145.05) -- (240.02,129.33) ; \draw [shift={(238.64,127.88)},
  rotate = 406.4] [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw
  opacity=1 ][line width=0.75] (6.56,-1.97) .. controls (4.17,-0.84)
and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84)
.. (6.56,1.97) ;
%Straight Lines [id:da4986041081219512] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (248.67,152.06) -- (237.99,143.3) ; \draw [shift={(236.44,142.03)},
  rotate = 399.35] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da8801881465271447] 
\draw [color={rgb, 255:red, 0; green, 13; blue, 255 } ,draw opacity=1
] (276.23,135.96) -- (277.24,114.27) ; \draw [shift={(277.33,112.28)},
  rotate = 452.65] [color={rgb, 255:red, 0; green, 13; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da03421807598256077] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (266.8,136.24) -- (265.75,122.47) ; \draw [shift={(265.59,120.47)},
  rotate = 445.6] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da42029107361898876] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(297.28,170.57) -- (318.97,171.32) ; \draw [shift={(320.97,171.39)},
  rotate = 181.97] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da34990856882657484] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (296.89,161.15) -- (310.64,159.92) ; \draw [shift={(312.64,159.75)},
  rotate = 534.9200000000001] [color={rgb, 255:red, 65; green, 117;
    blue, 5 } ,draw opacity=1 ][line width=0.75] (6.56,-1.97)
.. controls (4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls
(1.99,0.18) and (4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da8568463836228528] 
\draw [color={rgb, 255:red, 0; green, 13; blue, 255 } ,draw opacity=1
] (219.17,206) -- (205.78,215.82) ; \draw [shift={(204.17,217)},
  rotate = 323.75] [color={rgb, 255:red, 0; green, 13; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da6974865872969185] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (233.08,216.87) -- (226.48,224.49) ; \draw [shift={(225.17,226)},
  rotate = 310.94] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da9473192312247024] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(319.86,212.07) -- (329.76,222.08) ; \draw [shift={(331.17,223.5)},
  rotate = 225.31] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da5546786921432448] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (330.42,198.47) -- (338.39,202.58) ; \draw [shift={(340.17,203.5)},
  rotate = 207.31] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da2980392776012871] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(208.09,160.96) -- (192.16,160.11) ; \draw [shift={(190.17,160)},
  rotate = 363.07] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da3672184968558653] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (208,173.95) -- (198.17,173.99) ; \draw [shift={(196.17,174)},
  rotate = 359.77] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da11017355529948814] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(325.09,129.6) -- (335.44,123.51) ; \draw [shift={(337.17,122.5)},
  rotate = 509.55] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da6955212381544827] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (312.53,116.67) -- (319.71,109.87) ; \draw [shift={(321.17,108.5)},
  rotate = 496.58] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da3603103334401869] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(232.17,118) -- (218.49,102.5) ; \draw [shift={(217.17,101)}, rotate =
  408.58000000000004] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da6369305975573354] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (220.59,128.62) -- (212.8,123.15) ; \draw [shift={(211.17,122)},
  rotate = 395.12] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da3610436111733121] 
\draw [color={rgb, 255:red, 0; green, 13; blue, 255 } ,draw opacity=1
] (278.17,102) -- (279.07,84) ; \draw [shift={(279.17,82)}, rotate =
  452.86] [color={rgb, 255:red, 0; green, 13; blue, 255 } ,draw
  opacity=1 ][line width=0.75] (6.56,-1.97) .. controls (4.17,-0.84)
and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84)
.. (6.56,1.97) ;
%Straight Lines [id:da8453237207742365] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (263.17,103) -- (262.39,95.99) ; \draw [shift={(262.17,94)}, rotate
  = 443.66] [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw
  opacity=1 ][line width=0.75] (6.56,-1.97) .. controls (4.17,-0.84)
and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84)
.. (6.56,1.97) ;
%Straight Lines [id:da8500846118815689] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(335.91,171.88) -- (347.17,172.41) ; \draw [shift={(349.17,172.5)},
  rotate = 182.66] [color={rgb, 255:red, 4; green, 0; blue, 255 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da35490233668558013] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (337.69,158.9) -- (345.19,157.79) ; \draw [shift={(347.17,157.5)},
  rotate = 531.61] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;
%Straight Lines [id:da09061677521708489] 
\draw [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw opacity=1 ]
(270.17,232) -- (269.63,246.1) ; \draw [shift={(269.55,248.1)}, rotate
  = 272.19] [color={rgb, 255:red, 4; green, 0; blue, 255 } ,draw
  opacity=1 ][line width=0.75] (6.56,-1.97) .. controls (4.17,-0.84)
and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84)
.. (6.56,1.97) ;
%Straight Lines [id:da33913903399046463] 
\draw [color={rgb, 255:red, 65; green, 117; blue, 5 } ,draw opacity=1
] (285.64,228.91) -- (287.72,238.05) ; \draw [shift={(288.17,240)},
  rotate = 257.17] [color={rgb, 255:red, 65; green, 117; blue, 5 }
  ,draw opacity=1 ][line width=0.75] (6.56,-1.97) .. controls
(4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and
(4.17,0.84) .. (6.56,1.97) ;

% Text Node
\draw (204,138.4) node [anchor=north west][inner sep=0.75pt] {$\Sigma
  $};
% Text Node
\draw (273.23,157.14) node  [font=\scriptsize]  {$\boldsymbol{Q}$};

\end{tikzpicture}