### 2021-08-17 polaritonesferico.code.tex

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
%Shape: Circle [id:dp14966767726976038] 
\draw [fill={rgb, 255:red, 248; green, 231; blue, 28 } ,fill opacity=1
] (219.75,132.08) .. controls (219.75,101.2) and (244.78,76.17)
.. (275.67,76.17) .. controls (306.55,76.17) and (331.58,101.2)
.. (331.58,132.08) .. controls (331.58,162.97) and (306.55,188)
.. (275.67,188) .. controls (244.78,188) and (219.75,162.97)
.. (219.75,132.08) -- cycle ;
%Shape: Circle [id:dp9677285416132982] 
\draw [fill={rgb, 255:red, 248; green, 231; blue, 28 } ,fill opacity=1
][dash pattern={on 4.5pt off 4.5pt}] (219.75,108.08) .. controls
(219.75,77.2) and (244.78,52.17) .. (275.67,52.17) .. controls
(306.55,52.17) and (331.58,77.2) .. (331.58,108.08) .. controls
(331.58,138.97) and (306.55,164) .. (275.67,164) .. controls
(244.78,164) and (219.75,138.97) .. (219.75,108.08) -- cycle ;
%Shape: Path Data [id:dp41415328569096044] 
\draw [draw opacity=0][fill={rgb, 255:red, 25; green, 23; blue, 2 }
  ,fill opacity=0.53 ] (275.67,164) .. controls (248.27,164) and
(225.42,145.14) .. (220.17,120.08) .. controls (225.42,95.03) and
(248.27,76.17) .. (275.67,76.17) .. controls (303.06,76.17) and
(325.91,95.03) .. (331.17,120.08) .. controls (325.91,145.14) and
(303.06,164) .. (275.67,164) -- cycle ;

%Straight Lines [id:da9936815684711549] 
\draw [line width=1.5] (273.03,112.79) -- (273.17,126.96) ; \draw
      [shift={(273,109.79)}, rotate = 89.44] [color={rgb, 255:red, 0;
          green, 0; blue, 0 } ][line width=1.5] (8.53,-2.57)
      .. controls (5.42,-1.09) and (2.58,-0.23) .. (0,0) .. controls
      (2.58,0.23) and (5.42,1.09) .. (8.53,2.57) ;
%Straight Lines [id:da577848764958209] 
\draw [color={rgb, 255:red, 0; green, 4; blue, 243 } ,draw opacity=1
][line width=1.5] (235.87,105.83) -- (236.25,134.39) ; \draw
      [shift={(235.83,102.83)}, rotate = 89.24] [color={rgb, 255:red,
          0; green, 4; blue, 243 } ,draw opacity=1 ][line width=1.5]
      (8.53,-2.57) .. controls (5.42,-1.09) and (2.58,-0.23) .. (0,0)
      .. controls (2.58,0.23) and (5.42,1.09) .. (8.53,2.57) ;
%Straight Lines [id:da23709733272432076] 
\draw [color={rgb, 255:red, 40; green, 0; blue, 255 } ,draw opacity=1
][line width=1.5] (312.79,105.99) -- (313.17,134.55) ; \draw
      [shift={(312.75,102.99)}, rotate = 89.24] [color={rgb, 255:red,
          40; green, 0; blue, 255 } ,draw opacity=1 ][line width=1.5]
      (8.53,-2.57) .. controls (5.42,-1.09) and (2.58,-0.23) .. (0,0)
      .. controls (2.58,0.23) and (5.42,1.09) .. (8.53,2.57) ;
%Straight Lines [id:da356188359843302] 
\draw [color={rgb, 255:red, 22; green, 0; blue, 255 } ,draw opacity=1
][line width=1.5] (256.85,106.76) -- (257,135.32) ; \draw
      [shift={(256.83,103.76)}, rotate = 89.7] [color={rgb, 255:red,
          22; green, 0; blue, 255 } ,draw opacity=1 ][line width=1.5]
      (8.53,-2.57) .. controls (5.42,-1.09) and (2.58,-0.23) .. (0,0)
      .. controls (2.58,0.23) and (5.42,1.09) .. (8.53,2.57) ;
%Straight Lines [id:da517585674881536] 
\draw [color={rgb, 255:red, 22; green, 0; blue, 255 } ,draw opacity=1
][line width=1.5] (292.02,106.91) -- (292.17,135.47) ; \draw
      [shift={(292,103.91)}, rotate = 89.7] [color={rgb, 255:red, 22;
          green, 0; blue, 255 } ,draw opacity=1 ][line width=1.5]
      (8.53,-2.57) .. controls (5.42,-1.09) and (2.58,-0.23) .. (0,0)
      .. controls (2.58,0.23) and (5.42,1.09) .. (8.53,2.57) ;
%Straight Lines [id:da6814351373246353] 
\draw [color={rgb, 255:red, 126; green, 211; blue, 33 } ,draw
  opacity=1 ][line width=0.75] (283.14,134.98) -- (282.7,105.43) ;
\draw [shift={(283.17,136.98)}, rotate = 269.15] [color={rgb, 255:red,
    126; green, 211; blue, 33 } ,draw opacity=1 ][line width=0.75]
(8.74,-2.63) .. controls (5.56,-1.12) and (2.65,-0.24) .. (0,0)
.. controls (2.65,0.24) and (5.56,1.12) .. (8.74,2.63) ;
%Straight Lines [id:da19196139788350697] 
\draw [color={rgb, 255:red, 126; green, 211; blue, 33 } ,draw
  opacity=1 ][line width=0.75] (250.16,135.21) -- (249.72,105.65) ;
\draw [shift={(250.19,137.21)}, rotate = 269.15] [color={rgb, 255:red,
    126; green, 211; blue, 33 } ,draw opacity=1 ][line width=0.75]
(8.74,-2.63) .. controls (5.56,-1.12) and (2.65,-0.24) .. (0,0)
.. controls (2.65,0.24) and (5.56,1.12) .. (8.74,2.63) ;
%Straight Lines [id:da7557269004865166] 
\draw [color={rgb, 255:red, 126; green, 211; blue, 33 } ,draw
  opacity=1 ][line width=0.75] (299.33,134.18) -- (299.06,104.62) ;
\draw [shift={(299.35,136.18)}, rotate = 269.48] [color={rgb, 255:red,
    126; green, 211; blue, 33 } ,draw opacity=1 ][line width=0.75]
(8.74,-2.63) .. controls (5.56,-1.12) and (2.65,-0.24) .. (0,0)
.. controls (2.65,0.24) and (5.56,1.12) .. (8.74,2.63) ;
%Straight Lines [id:da6683208735835106] 
\draw [color={rgb, 255:red, 126; green, 211; blue, 33 } ,draw
  opacity=1 ][line width=0.75] (266.08,135.02) -- (265.81,105.47) ;
\draw [shift={(266.1,137.02)}, rotate = 269.48] [color={rgb, 255:red,
    126; green, 211; blue, 33 } ,draw opacity=1 ][line width=0.75]
(8.74,-2.63) .. controls (5.56,-1.12) and (2.65,-0.24) .. (0,0)
.. controls (2.65,0.24) and (5.56,1.12) .. (8.74,2.63) ;

% Text Node
\draw (268,47.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (285,52.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (298,58.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (307,66.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (314,79.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (232,66.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (226,76.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (240,57.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (252,51.4) node [anchor=north west][inner sep=0.75pt] {$-$};
% Text Node
\draw (222,139.4) node [anchor=north west][inner sep=0.75pt] {$+$};
% Text Node
\draw (231,153.4) node [anchor=north west][inner sep=0.75pt] {$+$};
% Text Node
\draw (241,163.4) node [anchor=north west][inner sep=0.75pt] {$+$};
% Text Node
\draw (253,168.4) node [anchor=north west][inner sep=0.75pt] {$+$};
% Text Node
\draw (284.32,167.86) node [anchor=north west][inner sep=0.75pt]
      [rotate=-359.43] {$+$};
% Text Node
\draw (298.37,159.96) node [anchor=north west][inner sep=0.75pt]
      [rotate=-359.37] {$+$};
% Text Node
\draw (309.32,148.14) node [anchor=north west][inner sep=0.75pt]
      [rotate=-358.41] {$+$};
% Text Node
\draw (316.41,137.18) node [anchor=north west][inner sep=0.75pt]
      [rotate=-358.48] {$+$};
% Text Node
\draw (268,169.4) node [anchor=north west][inner sep=0.75pt] {$+$};
% Text Node
\draw (273,119.4) node [anchor=north west][inner sep=0.75pt]
      [font=\tiny] {$\vec{\mathbf{\zeta }}$};

\end{tikzpicture}